---
title: "Project 1"
subtitle: "Reproductive Research using RMarkdown"
author: "Huiyue Li, Ying Gao, & Mian Wei<- Huiyue Li"
output: word_document
---
-----
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction
This exercise will use the **esoph** dataset, which is part of base R. The *esoph* dataset contains data from a case-control study of esophageal cancer in Ille-et-Vilaine, France.  
  
The *esoph* dataset has `r nrow(esoph)` rows and `r ncol(esoph)` variables, including:  
1. Age group  
2. Alcohol consumption  
3. Tobacco consumption  
4. Number of cases  
5. Number of controls  

## Data Details
* This dataset is aggregated, which means each row represents one distinct group that includes the number of cases and controls in each group. 
* There are six age groups:  
    + 25-34 
    + 35-44
    + 45-54 
    + 55-64 
    + 65-74 
    + 75+ 
* There are four alcohol consumption groups:
    + 0-39 *g/day*
    + 40-79 *g/day*
    + 80-119 *g/day*
    + 120+ *g/day*
* There are four tobacco consumption groups:
    + 0-9 *g/day*
    + 10-19 *g/day*
    + 20-29 *g/day*
    + 30+ *g/day*

## Data Processing
We want to know the percent of cases by exposure group. The code below calculates the total number of participants and the percentages for cases.
```{r, echo=TRUE, message=FALSE}
library(dplyr)

#create new variables
esoph$total = esoph$ncases + esoph$ncontrols
esoph$cases_p = esoph$ncases/esoph$total

#calculate mean percent of cases by age group
cases_by_age = esoph %>% group_by(agegp) %>% summarize(mean=100*mean(cases_p))
```
## Results 
### Table
Create a table showing the mean percentage of cases by age group.
```{r,echo=FALSE}
library(knitr)
kable(cases_by_age,col.names = c("Age Group","Percent Cases"),caption = "Table 1: Percent Cases by Age", digits = 2)
```
### Figure
The figure below shows the relationship between tobacco usage and esophageal cancer.
```{r,echo=TRUE,message=FALSE,fig.asp=2/3}
#code to create side-by-side boxplots

library(ggplot2)
library(wesanderson)

colors = wes_palette("Royal1", 4, type = c("discrete"))
ggplot(data=esoph, aes(x=tobgp, y=cases_p)) + geom_boxplot(fill=colors) + ggtitle("Distribution of esophageal cancer by tobacco use") + xlab("Tobacco Use (g/day)") + ylab("Percent Cases")
```

