---
title: "Project 2"
author: "Huiyue Li, Yujia Wang, Lujun Zhang<- Huiyue Li"
output:
  word_document: default
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,loadpackage, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```

### Import both datasets
```{r, data, message=FALSE}
# input the two datasets
encounter<- read_csv("Encounter Level Data.csv")
patient <- read_csv("Patient Level Data.csv")
```
<br>

### Merge the patient level data into the encounter level data
```{r,merge, message=FALSE}
# before merging the two datasets, we need to identify the key variable unique in these two tibbles
# look at the two datasets firstly
head(encounter)
head(patient)
# from the result, we can identify the variable MRN is the key variable
# merge the patient data into the encounter data using the key variable
encounter_patient<-merge(encounter,patient,by=c("MRN"))
```
<br>

### Re-categorize WBC into a categorical variable
```{r, re-categorize, message=FALSE}
# categorize WBC
WBC1<-"Not Taken"
WBC1[encounter_patient$WBC<3.2]="Low"
WBC1[3.2<=encounter_patient$WBC & encounter_patient$WBC<=9.8]="Normal"
WBC1[9.8<encounter_patient$WBC]="High"
encounter_patient$WBC<-WBC1
# we still have NA in WBC, then turn the NA into "Not Taken"
# use which[] to identify the position of NA in WBC
NT=which(is.na(encounter_patient$WBC))
# replace the NA with "Not Taken"
encounter_patient$WBC[NT]="Not Taken"
```
<br>

### print a table of the categorical WBC variable
```{r, table}
# obtain the dataframes of the counts and percentages for each category
count<-data.frame(table(encounter_patient$WBC))
proportion<-data.frame(prop.table(table(encounter_patient$WBC))*100)
# combine dataframes and print the table
WBC_table<-merge(count,proportion,by=c("Var1"))
kable(WBC_table,col.names = c("WBC","Count","Percentage (%)"),caption = "Table1: for categorical WBC variable",digits = 3,align = "c")
```
<br>
 
### Calculate & print a table of the mean BMI for the following MRNs: CI6950, IW9164, HJ8458, & XE4615
```{r, table2, warning=FALSE,message=FALSE}
MRNs<-encounter_patient%>%
# filter the rows required
  filter(MRN=="CI6950"|MRN=="IW9164"|MRN=="HJ8458"|MRN=="XE4615")%>% 
  group_by(MRN)%>%
  summarise(mean=mean(BMI,na.rm = T),.groups="drop_last")
# print the table
kable(MRNs,col.names = c("MRN","Mean of BMI"),caption = "Table 2: the mean of BMI for the MRN",digits = 3,align = "c")
```
<br>

### Create a table showing how many hospital encounters occurred each year
```{r, table3, message=FALSE}
library(lubridate)
# convert the date into the standard format, and add the year to a new column of the dataframe
Date<-as.Date(encounter_patient$contact_date,format = '%m/%d/%Y')
encounter_patient$Year<-year(Date)
# print the table
Year_en<-encounter_patient%>%
  filter(encounter_patient$enc_type=="Hospital Encounter")%>%
  group_by(Year)%>%
  summarise(n=n(),.groups="drop_last")
kable(Year_en, col.names = c("Year","Count"),caption = "Table 3: the number of hospital encounters each year",align = "c")

```
<br>

### Create & print a table of the counts & percentages of race, financial class, hypertension, congestive heart failure, and diabetes
```{r,final table,message=FALSE}
# since the five variables are all from the 'patient' dataframe, then use the 'patient' dataframe to print table (to avoid the replicate rows)
# for race
race<-patient%>%
  group_by(race)%>%
  summarise(Count1=n(),.groups="drop_last")%>%
  mutate(Proportion1=prop.table(Count1)*100)
  colnames(race)<-c("**Classification of Race**","**Count**","**Percentage (%)**")

# for financial class
financial<-patient%>%
  group_by(financialclass)%>%
  summarise(Count2=n(),.groups="drop_last")%>%
  mutate(Proportion2=prop.table(Count2)*100)
 colnames(financial)<-c("**Classification of Financialclass**","**Count**","**Percentage (%)**")

# for hypertension
hyper<-patient%>%
  group_by(hypertension)%>%
  summarise(Count3=n(),.groups="drop_last")%>%
  mutate(Proportion3=prop.table(Count3)*100)
 colnames(hyper)<-c("**Classification of Hypertension**","**Count**","**Percentage (%)**")

# for congestive heart failure (CHF)
CHF<-patient%>%
  group_by(CHF)%>%
  summarise(Count4=n(),.groups="drop_last")%>%
  mutate(Proportion4=prop.table(Count4)*100)
 colnames(CHF)<-c("**Classification of CHF**","**Count**","**Percentage (%)**")

# for diabetes
diabetes<-patient%>%
  group_by(diabetes)%>%
  summarise(Count5=n(),.groups="drop_last")%>%
  mutate(Proportion5=prop.table(Count5)*100)
 colnames(diabetes)<-c("**Classification of Diabetes**","**Count**","**Percentage (%)**")

# print the table 
kable(list(race,financial,hyper,CHF,diabetes),caption = "Table 4: for the following five variables",align = "c",digits = 3)
  
```
<br>

### Create a histogram of the distress score
```{r, histogram, message=FALSE,warning=FALSE}
# create the histogram
qplot(encounter_patient$distress,main = "# Histogram of the distress score",geom = "histogram",binwidth=0.5,fill=I("rosybrown1"),col=I("cornsilk"),xlab = "distress score",ylab = "frequency")
```


