---
title: "Project 4"
author: "Huiyue Li, Xinyi Fan, Weiliang Tian, Yujia Wang<- Huiyue Li"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

```{r packages}
# load the packages we need to use
library(knitr)
library(here)

# using the here package 
here::here()
```
<br>

#### Write a function that takes two continuous variables, standardizes each of them, and calculates the Spearman correlation between them for both the un-standardized and standardized variables (the function should return both correlations)
```{r function}
# input the data set we require
dat<-read.csv("P4_dat.csv")
# check if there is any NA
sum(is.na(dat))
# Function of standardization and calculating the Spearman correlation
Corr<-function(x,y){
  x_stand=(x-mean(x))/sd(x)    #standardize x
  y_stand=(y-mean(y))/sd(y)    #standardize y
  cor_unstand=round(cor(x,y,method = "spearman"),3) #calculate the spearman correlation for un-standardized variables
  cor_stand=round(cor(x_stand,y_stand,method = "spearman"),3) #calculate the spearman correlation for standardized variables
  return(c(cor_unstand,cor_stand)) #return both correlations
}
```
<br>

#### Create a table showing the correlations for each pair 
```{r table}
# obtain the number of column in the data set
k=ncol(dat)
#  create an empty vector for loop results
table_corr=NULL
# create a FOR loop to get the spearman correlation of each pair in the data set
for (i in 1:(k-1)) {
  for (j in (i+1):k) {
    table_corr=rbind(table_corr,c(paste("d",i," and d",j,sep = ""),Corr(dat[,i],dat[,j])))
  }
}
# display the result by kable
kable(table_corr,col.name=c("Pair","Un-standardized","Standardized"),caption="# Spearman correlation between each pair for both Un-standardized and Standardized variables",align="c")
```
<br>

#### Comment on the effect of standardization on Spearman correlation
Based on the result shown in the table, we can easily find that there is no difference between Spearman correlation for un-standardized and standardized variables, which means there is no effect of standardization on Spearman correlation.

